### 2018-05-15 Paper/method.tex

In this section, we outline the underlying framework for ``adaptively" constructing a 
reduced-order surrogate (ROS) using sensitivity analysis.  The proposed methodology
is described as adaptive since the ROS is constructed only in situations where
it is expected to yield computational dividend as discussed further below. 
The term `reduced-order' implies that the surrogate is constructed in a 
subspace that sufficiently captures the uncertainty in the model output. 
In the context of PC surrogates used in this work, it implies that the dimensionality
of the polynomial basis functions is effectively reduced.

Dimension-reduction is guided by global sensitivity analysis (GSA) of the uncertain model parameters. 
A typical approach to GSA involves estimating the Sobol sensitivity indices~\cite{Sobol:2001}
in the case of non-linear mathematical models. However, as discussed earlier in 
sections~\ref{sec:intro} and~\ref{sec:bg}, determining converged 
estimates of Sobol sensitivity indices typically requires tens of thousands
of model evaluations. Consequently, the exercise becomes prohibitive especially in
situations where the model runs are compute-intensive. In the proposed framework,
we adopt a novel approach for constructing an ROS by estimating the upper-bound 
($\widehat{\mathcal{C}_i\mu_i}$, see Eq.~\ref{eq:bound}) on Sobol total
effect index ($\mathcal{T}(\theta_i)$) for each parameter, $\theta_i$, and use it
as a metric to screen parameters that are relatively unimportant. A generalized
methodology for parameter screening is provided below in Algorithm~\ref{alg:screen}.

@@

In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically, we could use finite-difference
to approximate its value as shown in Eq.~\ref{eq:partial}. However, as discussed earlier
in section 2.1, estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires 
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in 
case the derivative is available analytically. Hence, computational effort when using finite differences
is expected to increase by a factor of $(N_p+1)$.

Based on the steps outlined in the above algorithm, it can be said that the set of sample points used for
estimating the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks
between successive iterations as well as a certain degree of convergence in estimates of the screening 
metric for each uncertain parameter is accomplished. However, the amount of computational effort associated
the screening process is limited by the choice of maximum number of iterations, $s_\text{max}$. Key inputs
include a limiting value of the maximum relative change in the sensitivity measure between succesive iterations
($\tau$), a limiting ratio ($\tau_\text{screen}$) of the sensitivity metric relative to its maximum value,
 a fraction $\beta$ to guide the number of new samples 
($ \lceil \beta n_1 \rceil$ denotes an integer $\leq~\beta n_1$) at each iteration,
a new set of samples ($\{ \bm{\theta}_k \}_{k = 1}^{n_0}$) for the initial screening step in the algorithm,
and the set of all the available gradient evaluations ($\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$). The outputs
include the set of active indices ($\mathcal{I}_\text{active}$) corresponding to the important parameters,
the total number of available model evaluations ($N_\text{total}$), and the enriched set of gradient evaluations
(($\{ \bm{g}^k \}_{k=1}^{N_\text{total}}$)). Parameter screening is an integral part of the overall framework,
illustrated using a flow-diagram in Figure~\ref{fig:flow}.

@@

We begin by setting aside computational resources for constructing a 
validation test suite to be used for assessing the accuracy of the resulting
surrogate. Obviously, the resources allocated for this purpose
depend upon the application as well as total amount of available resource. The applicability
of the proposed framework depends upon the choice of the model output. Since the screening metric
involves computation of partial derivatives in the full parameter space, the output must exhibit
a smooth dependence on each parameter for the approximation to be reasonable. It is therefore likely
that for a given application involving multiple outputs, the ROS can only be constructed for a 
selected few. Hence, where possible, it is prudent to assess the nature of the input-output
relationship for a given model prior to implementing the present framework. The set of required inputs
for parameter screening are initialized, and model evaluations at $n_0$ random samples in the full-space
are computed. These evaluations are used to construct a surrogate in the full-space (FSS) using regression-based
techniques. If the surrogate is found to be sufficiently accurate for the given application, the process
is terminated. However, it is highly unlikely that a full-space surrogate, constructed using a reasonably 
small number of model evaluations would result in a faithful representation of the input-output relationship. 
The available set of model evaluations are utlized and further enhanced during parameter screening as
discussed earlier. At the end of screening, the set of active indices, $\mathcal{I}_\text{active}$ is used
to evaluate $\alpha$ in the following equation:

\be
\alpha = \frac{|\mathcal{I}_\text{active}|}{N_p}
\label{eq:alpha}
\ee

\noindent where $|\mathcal{I}_\text{active}|$ denotes the cardinality of $\mathcal{I}_\text{active}$.
Scope for dimension-reduction increases as $\alpha$ decreases. Hence, if $\alpha$ is
considered to be small and computational gains are expected owing to dimension reduction, the ROS
is constructed and verified for accuracy using a combination of model evaluations used for screening
and those associated with the validation test suite. On the other hand, if $\alpha$ is considerably
large, the set of inputs required for screening are updated as needed, and a new set of $n_0$ samples
and corresponding model evaluations are generated. The FSS is re-constructed using the enriched set
of evaluations and the aforementioned analysis is repeated as illustrated in the flow-diagram.

The amount of computational effort associated with the presented methodology can be mainly attributed to
two steps: I. Parameter Screening, and II. Constructing a converged ROS. Computational gains are realized
in situations where constructing the surrogate in the full parameter space is more expensive than the combined cost
associated with these steps. Determining the optimal allocation of computational resources for these steps, however, 
is not possible a priori. Hence, in the proposed framework, we exploit the set of model evaluations used
in parameter screening to simultaneously construct the FSS while keeping a track of its accuracy using the
validation test suite. This would help address situations where significant dimension
reduction is not possible, and hence, constructing the ROS might result in a computational disadvantage. We
suggest using a small number of samples in the initial screening step (say, $n_0$ = 5) and a relatively large
$\tau$ (say, $\mathcal{O}(10^{-1})$) as a starting point with possible reduction in $\tau$ during
subsequent screenings.

In order to assess convergence of the resulting surrogate, one could estimate the leave-one-out cross 
validation error as follows:

\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee

\noindent where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash i}}$
is the PC surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$ model realization. 
From Eq.~\ref{eq:loo}, it appears that $N_l$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.
However, in practice a modified formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009},
independent of $G^{\mbox{\tiny {PC}\textbackslash i}}$ is used. Accuracy of the
surrogate could be assessed by evaluating the relative L-2 norm of the difference in predictions
between the model and the surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:

\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee

\noindent where $N_v$ is the number of sampling points in the full parameter space at which model
evaluations are available. Accuracy of the surrogate could be further investigated by comparing
probability density functions (PDFs) of the model output based on model evaluations in the full
parameter space and surrogate predictions corresponding to a large number of samples.

>>>>

In this section, we outline the underlying framework for \emph{adaptively}
constructing a reduced-order surrogate (ROS) using sensitivity analysis.  The
proposed methodology is described as adaptive since the ROS is constructed only
in situations where it is expected to yield computational dividend as discussed
further below.  The term reduced-order implies that the surrogate is
constructed in a subspace that sufficiently captures the uncertainty in the
model output.~\alennote{The next sentence seems unneeded, and 
might be confusing; consider removing it, or clarify that in 
the case of PCE, we construct PC bases as multivariate polynomials
on a parameter space with reduced dimension.}
In the context of PC surrogates used in this work, it implies
that the dimensionality of the polynomial basis functions is effectively
reduced.

Dimension reduction is guided by global sensitivity analysis (GSA).  A typical
approach to GSA involves estimating the Sobol sensitivity
indices~\cite{Sobol:2001} in the case of non-linear mathematical models.
However, as discussed earlier in sections~\ref{sec:intro} and~\ref{sec:bg},
determining accurate estimates of Sobol sensitivity indices typically requires
tens of thousands of model evaluations. Consequently, the exercise becomes
prohibitive for expensive-to-evaluate models.  In the proposed framework, we
adopt a novel approach for constructing an ROS based on estimating the
upper-bound $\widehat{\mathcal{C}_i\mu_i}$, given in~\eqref{eq:bound}, on 
Sobol total effect index ($\mathcal{T}(\theta_i)$) for each 
parameter, $\theta_i$, and use it as a metric to screen 
parameters that are relatively unimportant. A general
methodology for parameter screening is provided below in
Algorithm~\ref{alg:screen}.

@@

In situations where $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically,
we could use finite-difference to approximate its value as shown in
Eq.~\ref{eq:partial}. However, as discussed earlier in section 2.1,
~\alennote{Don't hard code section numbers, use labels, and refer to them.}
estimating $\bm{\mu}$  using $M$ samples in the full parameter space requires
model evaluations at $(N_p+1)M$ samples, compared to only $M$ samples in case
the derivative is available analytically. Hence, computational effort when
using finite differences is expected to increase by a factor of $(N_p+1)$.
Another possible approach, suitable in certain cases, is the use of
adjoint-based gradient computation. In such cases, each gradient computation,
which is more efficient than finite-differences---each gradient
evaluation requires a forward model solve and adjoint solve---but 
requires the availability of adjoint solvers.
\alennote{I will put some references there ...}

Based on the steps outlined in the above algorithm, it can be said that the set
of sample points used for estimating the screening metric,
$\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks between
successive iterations as well as a certain degree of convergence in estimates
of the screening metric for each uncertain parameter is accomplished. However,
the amount of computational effort associated the screening process is limited
by the choice of maximum number of iterations, $s_\text{max}$. Key inputs
include a limiting value of the maximum relative change in the sensitivity
measure between succesive iterations $\tau$, a limiting ratio
$\tau_\text{screen}$ of the sensitivity metric relative to its maximum value,
a fraction $\beta$ to guide the number of new samples $ \lceil \beta n_1
\rceil$ denotes an integer $\leq~\beta n_1$ at each iteration, a set of
samples $\{ \bm{\theta}_k \}_{k = 1}^{n_0}$ for the initial screening step in
the algorithm, and the set of available gradient evaluations $\{
\bm{g}^k \}_{k=1}^{n_0}$. The outputs include the set of active
indices $\mathcal{I}_\text{active}$ corresponding to the 
\emph{important parameters}, the total number of available model 
evaluations $N_\text{total}$, and the enriched set of gradient evaluations 
$\{ \bm{g}^k
\}_{k=1}^{N_\text{total}}$. Parameter screening is an integral part of the
overall framework, illustrated using a flow-diagram in Figure~\ref{fig:flow}.

@@

We begin by setting aside computational resources for constructing a validation
test suite to be used for assessing the accuracy of the resulting surrogate.
Naturally, the resources allocated for this purpose depend upon the application
as well as total amount of available resources. 
%
%The applicability of the
%proposed framework depends upon the choice of the model output. 
%
Since the
screening metric involves computation of partial derivatives in the full
parameter space, the output must exhibit differentiable 
dependence on each parameter. It is therefore likely that for a given
application involving multiple outputs, the ROS can only be constructed for a
selected few. Hence, where possible, it is important to assess the nature of the
input-output relationship for a given model prior to implementing the present
framework. The set of required inputs for parameter screening are initialized,
and model evaluations at $n_0$ random samples in the full-space are computed.
These evaluations are used to construct a surrogate in the full-space (FSS)
using regression-based techniques. If the surrogate is found to be sufficiently
accurate for the given application, the process is terminated. However, it is
likely that a full-space surrogate constructed using a 
small number of model evaluations would not provide a 
faithful representation of the input-output relationship.

The available set of model evaluations are utlized and further enhanced during
parameter screening as discussed earlier.  At the end of screening, the set of
active indices, $\mathcal{I}_\text{active}$, is used to evaluate $\alpha$ in the
following equation: 
\be
\alpha = \frac{|\mathcal{I}_\text{active}|}{N_p},
\label{eq:alpha}
\ee
\alennote{I noticed in the latex code the use of the \textbf{noindent} command;
please remove those, we will use latext setting to formation paragraphs.} where
$|\mathcal{I}_\text{active}|$ denotes the cardinality of
$\mathcal{I}_\text{active}$.  Scope for dimension-reduction increases as
$\alpha$ decreases.  Hence, if $\alpha$ is considered to be small and
computational gains are expected owing to dimension reduction, the ROS is
constructed and verified for accuracy using a combination of model evaluations
used for screening and those associated with the validation test suite.  On the
other hand, if $\alpha$ is considerably large (~\alennote{You mean close to
one?}), the set of inputs required for screening are updated as needed, and a
new set of $n_0$ samples and corresponding model evaluations are generated. The
FSS is re-constructed using the enriched set of evaluations and the
aforementioned analysis is repeated as illustrated in the flow-diagram.

The amount of computational effort associated with the presented methodology
can be mainly attributed to two steps: I.~Parameter Screening, and 
II.~Constructing a converged ROS. Computational gains are realized in situations
where constructing the surrogate in the full parameter space is more expensive
than the combined cost associated with these steps. Determining the optimal
allocation of computational resources for these steps, however, is not possible
a priori. Hence, in the proposed framework, we exploit the set of model
evaluations used in parameter screening to simultaneously construct the FSS
while keeping a track of its accuracy using the validation test suite. This
would help address situations where significant dimension reduction is not
possible, and hence, constructing the ROS might result in a computational
disadvantage. We suggest using a small number of samples in the initial
screening step (say, $n_0$ = 5) and a relatively large $\tau$ (say,
$\mathcal{O}(10^{-1})$) as a starting point with possible reduction in $\tau$
during subsequent screenings.

To assess convergence of the resulting surrogate, one could estimate the
leave-one-out cross validation error as follows:
\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee
where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N_l}\sum\limits_{i=1}^{N_l} G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny
{PC}\textbackslash i}}$ is the PC surrogate constructed using all but the
$i^{\mbox{\tiny{th}}}$ model realization.  From~\eqref{eq:loo}, it appears
that $N_l$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.  However,
in practice a modified formulation for $\epsilon_{\mbox{\tiny
LOO}}$~\cite{Blatman:2009}, independent of $G^{\mbox{\tiny {PC}\textbackslash
i}}$ is used. Accuracy of the surrogate could be assessed by evaluating the
relative $\ell_2$-norm of the difference in predictions between the model and the
surrogate ($\epsilon_{\mbox{\tiny{L-2}}}$), as follows:
\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee
where $N_v$ is the number of sampling points in the full parameter space at
which model evaluations are available. Accuracy of the surrogate could be
further investigated by comparing probability density functions (PDFs) of the
model output based on model evaluations in the full parameter space and
surrogate predictions corresponding to a large number of samples.
